Predicting Remaining Useful Life of NASA Aircraft Engines

### 2025-02-24 index.qmd
Added equation

While health care is the dominant field in which CPH is used, a study in
2021 provides an example of using CPH in insurance fields. An insurance
company in the Czech Republic provides policies for certain illness
events until the event occurs when the insured will receive money and
their contract is complete. To assist the company with predicting if an
individual will have one of the illness events occur, a model using CPH
was generated using gender, age, and region as the
covariates[@zapletal2021]. A study in 2023 focused on the length of
productive life of female Floridian goats to determine the factors that
would impact the longevity of the goats. CPH was used to determine that
the age of first kidding and the specific herd impacts the length of
productive life of the goats [@ziadi2023]. Both of these studies provide
examples of the reaches of CPH models outside the health industry.

@@

-   Detail the models or algorithms used.

-   Justify your choices based on the problem and data.

*The common non-parametric regression model is*
$Y_i = m(X_i) + \varepsilon_i$*, where* $Y_i$ *can be defined as the sum
of the regression function value* $m(x)$ *for* $X_i$*. Here* $m(x)$ *is
unknown and* $\varepsilon_i$ *some errors. With the help of this
definition, we can create the estimation for local averaging i.e.*
$m(x)$ *can be estimated with the product of* $Y_i$ *average and* $X_i$
*is near to* $x$*. In other words, this means that we are discovering
the line through the data points with the help of surrounding data
points. The estimation formula is printed below [@R-base]:*

$$
M_n(x) = \sum_{i=1}^{n} W_n (X_i) Y_i  \tag{1}
$$$W_n(x)$ *is the sum of weights that belongs to all real numbers.
Weights are positive numbers and small if* $X_i$ *is far from* $x$*.*

*Another equation:*

$$
y_i = \beta_0 + \beta_1 X_1 +\varepsilon_i
$$

@@

We will be using the test_FD002 and train_FD002 datasets available for
download at [NASA's Open Data
Portal](https://data.nasa.gov/Aerospace/CMAPSS-Jet-Engine-Simulated-Data/ff5v-kuh6/about_data)
and also available on the home page of our [GitHub
repository](https://github.com/JR-87/capstone4ds_reed).

```{r}
testurl <- 'https://raw.githubusercontent.com/JR-87/capstone4ds_reed/refs/heads/main/test_FD002.txt'  
test <- read.table(testurl, header = TRUE, sep = '')
head(test)
```

```{r}
trainurl <- 'https://raw.githubusercontent.com/JR-87/capstone4ds_reed/refs/heads/main/train_FD002.txt'  
test <- read.table(trainurl, header = TRUE, sep = '') 
head(test)
```

##

@@

@@

### Conclusion

>>>>

While health care is the dominant field in which CPH is used, a study in
2021 provides an example of using CPH in insurance fields. An insurance
company in the Czech Republic provides policies for certain illness
events until the event occurs when the insured will receive money and
their contract is complete. To assist the company with predicting if an
individual will have one of the illness events occur, a model using CPH
was generated using gender, age, and region as the covariates
[@zapletal2021]. A study in 2023 focused on the length of productive
life of female Floridian goats to determine the factors that would
impact the longevity of the goats. CPH was used to determine that the
age of first kidding and the specific herd impacts the length of
productive life of the goats [@ziadi2023]. Both of these studies provide
examples of the reaches of CPH models outside the health industry.

@@

*Notes:*

*How coxph was created and creator what it was made for. it is semi
parametric regression*

*explain mathematically with equation*

*How data needs to be formatted - time + event and covariates*

*how it works*

*assumptions - how to test assumptions*

-   *Detail the models or algorithms used.*

-   *Justify your choices based on the problem and data.*

CPH can be estimated as

$$
h(t|\mathbf{Z}) = h_0(t)\text{exp}(\Sigma_{k=1}^{p} \beta_kZ_k)
$$ {#eq-hazard}

where,

-   $h(t|\mathbf{Z})$ is the hazard rate at time $t$ for an individual
    with risk vector $\mathbf{Z}$
-   $h_0(t)$ is an arbitrary baseline hazard rate
-   $\mathbf{Z}$ is the set of $p$ covariates
-   $\beta$ is the set of $p$ coefficients that measure the impact of
    the covariates

@@

Each engine has an unknown amount of wear, manufacturing variation, and
sensor noise all of which impact the survival time. In the training
dataset, each engine is operating normally at the beginning of the time
series, develops a fault, and experiences failure at the end of the time
series. The conditions the engines are subjected to are represented by
three operational setting fields and 21 sensor measurement fields. In
the testing data, the engines have not experienced failure
[@NASA_CMAPSS].

```{r, echo = FALSE}
#packages
library(tidyverse)
library(gridExtra)
library(survival)
library(survminer)

```{r}
testurl <- 'https://raw.githubusercontent.com/JR-87/capstone4ds_reed/refs/heads/main/test_FD002.txt'

test <- read.table(testurl, header = TRUE, sep = '')
head(test)
```

```{r}
trainurl <- 'https://raw.githubusercontent.com/JR-87/capstone4ds_reed/refs/heads/main/train_FD002.txt'

train <- read.table(trainurl, header = TRUE, sep = '')
head(test)
```

The visual below shows a histogram and a box plot representing
distributions of survival times of the 260 unique engines in the
training data.

```{r, echo=FALSE}
distributions <- train %>% 
  group_by(id) %>%
  summarise(time = max(time))

Histogram <- ggplot(data = distributions,
       aes(x = time)) +
  geom_histogram(binwidth = 5,
                 fill = "grey")+
  labs(x = "Survival Time",
       y = "Frequency")+
  theme_bw()

box <- ggplot(data = distributions,
                  aes(y = time))+
  geom_boxplot(fill = "gray",
               color = "black")+
  theme_bw()

grid.arrange(Histogram,box,ncol = 2, top = "Distribution of Engine Longevity in Training Data")

```{r}
summary(distributions$time)

sd(distributions$time)
```

The engines have a median survival time of 199 iterations with a
standard deviation of 46.8. The longest surviving engine lasted 378
iterations, with the shortest survival time for all engines being 128.

@@

### Modeling

@@

### Results

## Conclusion
